\Huge Conferences in the Era of Expensive Carbon

### 2019-09-20 main.tex
Some BCP comments

How should ACM respond to these facts?  Two years ago, ACM's Special
Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
Committee to research this question~\cite{ClimateCommittee}.  After
investigating many options~\cite{ClimateCommitteReport} and discussing
intensively with members, conference organizers, ACM staff, and leaders of
other SIGs, we are putting forward two concrete proposals. First, all ACM
conferences should publicly {\em account} for the {\COtwoE} emitted as a
result of putting them on. Second, ACM \iflater\bcp{Are we really
  talking about ACM, in the first instance?  Or is the proposal really that
  individual SIGs should do this, until there is enough buy-in to do it at
  the ACM level?  The first is easier to say, but the second is the only
  plausible implementation path.  If there's space for at least a sentence
  or two on this (and other practicalities of implementation), it might be
  good.}\fi should put a \emph{price} on 
carbon in conference budgets, to create pressure on organizers to reduce
their footprints.

@@

Our first proposal is modest: {\bf Every ACM-sponsored conference should
publicly report its carbon footprint.} 
%
Public accounting will nudge conference organizers and attendees to change
their behavior. By analogy, chain restaurants in the US are required by law
to post calorie counts of food items on menu boards, and studies show that
enlightened customers order, on average, up to 50 fewer calories a day as a
result~\cite{menu}.  
%MWH: I don't know why this should be a parenthetical.  Why would the
%analogy be an aside? It is meant to connect the reader's understanding to
%what we are proposing, explciitly; it is meant to be read, not skipped
%over. 
% BCP: In general, parenthetical remarks are not meant to be skipped over:
% they are meant to be read as asides.  I think this reads better in parens,
% but I won't fight for it.

@@

Many organizations, including companies such as Google, Dell, Microsoft,
General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
and some academic societies, now buy carbon offsets in order to reduce their
net footprint. ACM conferences can do the same, and the purchases can be
mentioned in the public accounting we are proposing.  In fact, ACM recently
decided that conferences could offer attendees the option to pay for an
offset during registration.  Other scientific societies are going further,
rolling carbon offsets into the registration fees for all
participants.\iflater\bcp{In a later draft, we may be able to say that ICML,
NIPS, etc. are a case in point.  Ask Alexandre Lacoste about this.}\fi

@@

ACM should develop tools to gather and publicize this data. As a first step,
SIGPLAN recently built an air-travel-focused {\COtwoE} calculator for
conferences\iflater~\cite{carboncalculator}\bcp{Trim citation?  Is it even
  the right one? Link seems
  dead!}\fi. Users can upload conference registration data, and the
calculator will estimate the {\COtwoE} cost of air travel for those
sufficiently far away. This calculator could be expanded to account for
other forms of conference emissions.

@@

\begin{figure}[t]
\centering
\includegraphics[width=6in]{SIGPLAN-confs2.png}
\includegraphics[width=6in]{ParticipantsOriginAll2.png}
% \includegraphics[width=6in]{Distances.png}
\caption{Top: Carbon footprint per participant for travel to recent SIGPLAN
  conferences\iflater\bcp{Get better file from the other paper, or get
    better version of this one from Crista.}\fi.  The smallest dot (ICFP
  14, in Gothenburg, Sweden) represents .9 tons of {\COtwoE} per participant; the
  largest dot (ICFP 16, in Nara Japan) represents 1.94 tons per
  participant. Bottom: Breakdown of continent-of-origin for participants in
  each conference.  Colored bars represent percentages of participants whose
  home city is in each continent.
% Bottom: Distances traveled by participants, one way, in Km; middle dashed
% line is the median travel distance. 
}
\label{fig:confs}
\end{figure}

To illustrate the challenges, consider the problem of choosing a conference
location that minimizes emissions from participants' travel.  Using recent
registration data from four SIGPLAN conferences, Figure~\ref{fig:confs}
shows two ways of looking at the relation between locations and emissions. 
%
The top diagram shows an
estimated per-participant {\COtwoE} footprint for each instance of each
conference over the past ten years (excluding a few for which we
had trouble getting data), with larger dots representing more emissions.
Eyeballing this diagram, it seems that carbon-conscious 
organizers should hold all of these conferences in either the northeast US or
western Europe every year.  But the bottom visualization tells a
different story.  The horizontal colored bar at the bottom represents
the continent on which each conference was held, and each vertical bar gives
a breakdown of the participants in that conference, colored according to the
continent of their work address.  From a superficial glance
at the colors, it is clear that, though a minority come from far away, the
majority of participants in each conference are local to the region where the
conference is being held.\iflater\bcp{That point could be strengthened by adding some
  numbers (just in the text, not a figure) about what percentage of
  conference attendees attend even one additional instance of the same
  conference...}\fi{} This suggests that locating
conferences every year in the same one or two places would significantly
damage the 
research community by discouraging participation from other parts of the
world.  Indeed, one might conclude that, from the point of view of
strengthening the research community, conferences should move around as 
much as possible.
%
%
% \bcp{Explain the two diagrams.  Discuss how they can be used to draw
%   contradictory conclusions.  E.g., one of the many footprint-reduction
%   measures that have been discussed in SIGPLAN~\cite{ClimateCommitteReport}
%   is locating conferences near their historical ``center of mass.''  Looking
%   at the top diagram, one might thus be tempted to locate all conferences in
%   the northeast US or western Europe.  But the bottom figure shows that much
%   of the participation in each instance of a conference is local; always
%   holding a conference at its ``center of mass'' will damage local
%   communities everywhere else.  (Maybe this would be a good place to refer
%   to~\cite{ClimateCommitteReport}---we can enumerate some of the
%   alternatives that one might have hoped the data would help us choose
%   between.)}  
% %
% \bcp{Maybe we can replace the bottom part of the figure by some clearer way
%   of visualizing the observation that much of the attendance at each
%   conference is local?  E.g., what if, instead of color-coding the different
%   sections of the bars by continent, we instead color-coded them by distance
%   traveled (in buckets of 500 miles, say)?}
%
These disparate perspectives suggest that reducing conference emissions will
require genuinely painful compromises.  The impulse to avoid thinking about
it is entirely understandable!

>>>>

How should ACM respond to these facts?  Two years ago, ACM's Special
Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
Committee to research this question~\cite{ClimateCommittee}.  After
investigating many options~\cite{ClimateCommitteReport} and intensive and
wide-ranging discussions, we are putting forward two concrete
proposals. First, all ACM 
conferences should publicly {\em account} for the {\COtwoE} emitted as a
result of putting them on---in particular, from travel to
the conference. Second, ACM \iflater\bcp{Are we really
  talking about ACM, in the first instance?  Or is the proposal really that
  individual SIGs should do this, until there is enough buy-in to do it at
  the ACM level?  The first is easier to say, but the second is the only
  plausible implementation path.  If there's space for at least a sentence
  or two on this (and other practicalities of implementation), it might be
  good.}\fi should put a \emph{price} on 
carbon in conference budgets, to create pressure on organizers to reduce
their footprints.

@@

Our first proposal is modest: {\bf Every ACM-sponsored conference should
publicly report its carbon footprint}.  Critically, this accounting should include the emissions generated by participants' travel to and from each conference, as this will constitute the bulk of the conference's carbon footprint.
%
Public accounting will nudge conference organizers and attendees to change
their behavior. By analogy, chain restaurants in the US are required by law
to post calorie counts of food items on menu boards, and studies show that
enlightened customers order, on average, up to 50 fewer calories a day as a
result~\cite{menu}.  
%MWH: I don't know why this should be a parenthetical.  Why would the
%analogy be an aside? It is meant to connect the reader's understanding to
%what we are proposing, explciitly; it is meant to be read, not skipped
%over. 
% BCP: In general, parenthetical remarks are not meant to be skipped over:
% they are meant to be read as asides.  I think this reads better in parens,
% but I won't fight for it.

@@

Many organizations, including companies such as Google, Dell, Microsoft,
General Motors, Delta Airlines, Lyft, and Expedia, universities,
and academic societies, now buy carbon offsets in order to reduce their
net footprint. ACM conferences can do the same, and the purchases can be
mentioned in the public accounting we are proposing.  In fact, ACM recently
decided that conferences could offer attendees the option to pay for an
offset during registration.  Other scientific societies are going further,
rolling carbon offsets into the registration fees for all
participants.\iflater\bcp{In a later draft, we may be able to say that ICML,
NIPS, etc. are a case in point.  Ask Alexandre Lacoste about this.}\fi

@@

ACM should develop tools to gather and publicize this data. As a first step,
SIGPLAN recently built an air-travel-focused {\COtwoE} calculator for
conferences\iflater~\cite{carboncalculator}\bcp{Trim citation?  Is it even
  the right one? Link seems
  dead!  But Richard says ``As a reader, my first question is, how do I get
  access to this calculator? Can you cite this calculator in some way? It
  would be great to give a citation to where people can learn about it or
  use it, even if just, "To learn more about the calculator, e-mail
  _________."''}\fi. Users can upload conference registration data, and the
calculator will estimate the {\COtwoE} cost of air travel for those
sufficiently far away. This calculator could be expanded to account for
other forms of conference emissions.

@@

\begin{figure}[t]
\centering
\includegraphics[width=6in]{SIGPLAN-confs2.png}
\includegraphics[width=6in]{ParticipantsOriginAll2.png}
% \includegraphics[width=6in]{Distances.png}
\caption{Top: Carbon footprint per participant for travel to recent SIGPLAN
  conferences\iflater\bcp{Get better file from the other paper, or get
    better version of this one from Crista.}\fi.  The smallest dot (ICFP
  14, in Gothenburg, Sweden) represents .9 tons of {\COtwoE} per participant; the
  largest dot (ICFP 16, in Nara Japan) represents 1.94 tons per
  participant. Bottom: Breakdown of continent-of-origin for participants in
  each conference.  Colored bars represent percentages of participants whose
  home city is in each continent.  \iflater\bcp{Richard: I find this a
    little hard to follow. Maybe you could simply label the different groups
    of conferences by their location? I think the "(NA)", "(EU)", and
    "(IAS)" are a bit hard to read and decode at first, so that I didn't
    understand where each conference was located at first glance. (Adrianne
    had the same response when I showed her.) Maybe just put text under
    each, maybe with a horizontal bracket, that says "North America,"
    "Europe," and "Rest of world," with the bracket and/or words color-coded
    as well.}\fi 
% Bottom: Distances traveled by participants, one way, in Km; middle dashed
% line is the median travel distance. 
}
\label{fig:confs}
\end{figure}

\bcp{Richard: I get a little lost in this paragraph. Partly because I found the figure a bit confusing, partly because of the length of the paragraph. I think one fix might be to frontload your conclusion a bit with a topic sentence. E.g., start the paragraph with something like:
"For example, choosing a conference location with the goal of minimizing CO2e emissions from participants' travel results in trade-offs in who participates."}
To illustrate the challenges, consider the problem of choosing a conference
location that minimizes emissions from participants' travel.  Using recent
registration data from four SIGPLAN conferences, Figure~\ref{fig:confs}
shows two ways of looking at the relation between locations and emissions. 
%
The top diagram shows an
estimated per-participant {\COtwoE} footprint for each instance of each
conference over the past ten years (excluding a few for which we
had trouble getting data), with larger dots representing more emissions.
Eyeballing this diagram, it seems that carbon-conscious 
organizers should hold all of these conferences in either the northeast US or
western Europe every year.  But the bottom visualization tells a
different story.  The horizontal colored bar at the bottom represents
the continent on which each conference was held, and each vertical bar gives
a breakdown of the participants in that conference, colored according to the
continent of their work address.  From a superficial glance
at the colors, it is clear that, though a minority come from far away, the
majority of participants in each conference are local to the region where the
conference is being held.\iflater\bcp{That point could be strengthened by adding some
  numbers (just in the text, not a figure) about what percentage of
  conference attendees attend even one additional instance of the same
  conference...}\fi{} This suggests that locating
conferences every year in the same one or two places would significantly
damage the 
research community by discouraging participation from other parts of the
world.\bcp{Richard: The damage to the research community is a little unclear to me on first read. I would reword to something like "that always locating conferences in the same one or two places would significantly damage the heterogeneity of the research community by discouraging participation from other parts of the world". Or perhaps some better world, but more specificity of the damage.}  Indeed, one might conclude that, from the point of view of
strengthening the research community, conferences should move around as 
much as possible.
%
%
% \bcp{Explain the two diagrams.  Discuss how they can be used to draw
%   contradictory conclusions.  E.g., one of the many footprint-reduction
%   measures that have been discussed in SIGPLAN~\cite{ClimateCommitteReport}
%   is locating conferences near their historical ``center of mass.''  Looking
%   at the top diagram, one might thus be tempted to locate all conferences in
%   the northeast US or western Europe.  But the bottom figure shows that much
%   of the participation in each instance of a conference is local; always
%   holding a conference at its ``center of mass'' will damage local
%   communities everywhere else.  (Maybe this would be a good place to refer
%   to~\cite{ClimateCommitteReport}---we can enumerate some of the
%   alternatives that one might have hoped the data would help us choose
%   between.)}  
% %
% \bcp{Maybe we can replace the bottom part of the figure by some clearer way
%   of visualizing the observation that much of the attendance at each
%   conference is local?  E.g., what if, instead of color-coding the different
%   sections of the bars by continent, we instead color-coded them by distance
%   traveled (in buckets of 500 miles, say)?}
%
These disparate perspectives suggest that reducing conference emissions will
require genuinely painful compromises.  The impulse to avoid thinking about
it is entirely understandable!
